### 2012-10-02 problem_formulation.tex
more refactoring

\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes

 \title{An application of Graph Theory to MDAO problem formulation}

\author{
  David Pate, %
     \thanks{Georgia Tech}
  Dr. Brian German 
     \thanks{Georgia Tech...}
  Justin Gray,%
     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
 }

%\usepackage{setspace}
%\doublespace

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{appendix}
\usepackage[section]{placeins}

\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}

\bibliographystyle{aiaa}

\usepackage{lettrine}
\usepackage{verbatim}

%\usepackage{hyperref} %allows for the creation of actual text links
\begin{document}

\maketitle

\begin{abstract}
   blah blah blah
\end{abstract}

\section*{Nomenclature}

\begin{tabular}{l l} 
    AAO      & All-At- \\
    MDAO     & Multidisciplinary Design Analysis and Optimization \\
    FPF      & Fundamental Problem Formulation \\
\end{tabular}

\input{intro}
\input{graph_syntax}

\section{Example Problem}

\section{Applications}

\section{Conclusions}

\bibliography{library}
\end{document}