### 2012-10-02 XDSM/diagram_styles.tex
starting xdsm for simple

% Define all the styles used to produce XDSMs for MDO

% Component types
\tikzstyle{Optimization} = [rounded rectangle,draw,fill=blue!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{LP_Optimization} = [rectangle,draw,fill=blue!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Analysis} = [rectangle,draw,fill=green!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Function} = [rectangle,draw,fill=purple!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{MDA} = [rounded rectangle,draw,fill=orange!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Metamodel} = [rectangle,draw,fill=yellow!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{DOE} = [rounded rectangle,draw,fill=yellow!20,inner sep=6pt,minimum height=1cm,text badly centered]
%\tikzstyle{OptFunction} = [rectangle,draw,fill=red!20,inner sep=6pt,minimum height=1cm,text badly centered]

%% A simple command to give the repeated structure look for components and data
\tikzstyle{stack} = [double copy shadow]

%% A simple command to fade components and data, e.g. demonstrating a sequence of steps in an animation
\tikzstyle{faded} = [draw=black!50,fill=white,text opacity=0.5]

%% Simple fading commands for the lines
\tikzstyle{fadeddata} = [color=black!20]
\tikzstyle{fadedprocess} = [color=black!50]

% Component types for repeated structures (i.e. for parallel structures)
\tikzstyle{Optimization_i} = [double copy shadow, Optimization]
\tikzstyle{LP_Optimization_i} = [double copy shadow, LP_Optimization]
\tikzstyle{Analysis_i} = [double copy shadow, Analysis]
\tikzstyle{Function_i} = [double copy shadow, Function]
\tikzstyle{MDA_i} = [double copy shadow, MDA]
\tikzstyle{Metamodel_i} = [double copy shadow, Metamodel]
\tikzstyle{DOE_i} = [double copy shadow, DOE]

% Faded component types for, e.g. demonstrations of each step. We use these style definitions to "gray out" large parts of the diagram.
\tikzstyle{Optimization_fade} = [Optimization,fill=blue!10,draw=black!30,text opacity=0.3]
\tikzstyle{Analysis_fade} = [Analysis,fill=green!10,draw=black!30,text opacity=0.3]
\tikzstyle{Function_fade} = [Function,fill=purple!10,draw=black!30,text opacity=0.3]
\tikzstyle{MDA_fade} = [MDA,fill=orange!10,draw=black!30,text opacity=0.3]
\tikzstyle{Metamodel_fade} = [Metamodel,fill=yellow!10,draw=black!30,text opacity=0.3]
\tikzstyle{DOE_fade} = [DOE,fill=yellow!10,draw=black!30,text opacity=0.3]

\tikzstyle{Optimization_i_fade} = [Optimization_i,fill=blue!10,draw=black!30,text opacity=0.3]
\tikzstyle{Analysis_i_fade} = [Analysis_i,fill=green!10,draw=black!30,text opacity=0.3]
\tikzstyle{Function_i_fade} = [Function_i,fill=purple!10,draw=black!30,text opacity=0.3]
\tikzstyle{MDA_i_fade} = [MDA_i,fill=orange!10,draw=black!30,text opacity=0.3]
\tikzstyle{Metamodel_i_fade} = [Metamodel_i,fill=yellow!10,draw=black!30,text opacity=0.3]
\tikzstyle{DOE_i_fade} = [DOE_i,fill=yellow!10,draw=black!30,text opacity=0.3]

% Data types
\tikzstyle{DataInter} = [trapezium,trapezium left angle=75,trapezium right angle=105,draw,fill=black!10]
\tikzstyle{DataIO} = [trapezium,trapezium left angle=75,trapezium right angle=105,draw,fill=white]

% Data types for repeated structures
\tikzstyle{DataInter_i} = [double copy shadow, DataInter]
\tikzstyle{DataIO_i} = [double copy shadow, DataIO]

% Faded data types
\tikzstyle{DataInter_fade} = [DataInter,draw=black!30,fill=white,text opacity=0.3]
\tikzstyle{DataIO_fade} = [DataIO_i,draw=black!30,fill=white,text opacity=0.3]

\tikzstyle{DataInter_i_fade} = [DataInter_i,draw=black!30,fill=white,text opacity=0.3]
\tikzstyle{DataIO_i_fade} = [DataIO_i,draw=black!30,fill=white,text opacity=0.3]

% Edges
\tikzstyle{DataLine} = [color=black!40,line width=5pt]
\tikzstyle{ProcessHV} = [-,line width=1pt,to path={-| (\tikztotarget)}]
\tikzstyle{ProcessTip} = [-,line width=1pt]

% Faded edges
\tikzstyle{DataLine_fade} = [DataLine,color=black!10]
\tikzstyle{ProcessHV_fade} = [ProcessHV,color=black!30]
\tikzstyle{ProcessTip_fade} = [ProcessTip,color=black!30]

% Matrix options
\tikzstyle{MatrixSetup} = [row sep=3mm, column sep=2mm]

% Declare a background layer for showing node connections
\pgfdeclarelayer{data}
\pgfdeclarelayer{process}
\pgfsetlayers{data,process,main}

% A new command to split the component text over multiple lines
\newcommand{\MultilineComponent}[3]
{
	\begin{minipage}{#1}
	\begin{center}
		#2 \linebreak #3
	\end{center}
	\end{minipage}
}

% A new command to split the component text over multiple columns
\newcommand{\MultiColumnComponent}[5]
{
	\begin{minipage}{#1}
	\begin{center}
	#2 \linebreak #3
	\end{center}
	\begin{minipage}{0.49\textwidth}
	\begin{center}
	#4
	\end{center}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
	\begin{center}
		#5
	\end{center}
	\end{minipage}
	\end{minipage}
}