### 2013-04-18 problem_formulation.tex
working on formatting for SMO

\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes

 \title{A Graph Theory Based Approach to MDAO Problem Formulation}

\author{
  David Pate, %
  Justin Gray,%
  Dr. Brian German 
 }

 \institute { Justin Gray \at NASA Glenn Research Center, Mail Stop 5-11, 21000 Brookpark Rd Clevland OH 44135 %
  \and David Pate \at Gatech \and Dr. Brian German \at Gatech}

\journalname{Structrual and Multidisciplinary Optimization}

%\usepackage{setspace}
%\doublespace

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{appendix}
\usepackage[section]{placeins}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{booktabs}

\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}

\bibliographystyle{aiaa}

\usepackage{lettrine}
\usepackage{verbatim}

\newcommand{\txt}{\textrm}

%\usepackage{hyperref} %allows for the creation of actual text links
\begin{document}

\maketitle

\begin{abstract}
   blah blah blah
\end{abstract}

\section*{Nomenclature}

\begin{tabular}{l l} 
    AAO      & All-At- \\
    MDAO     & Multidisciplinary Design Analysis and Optimization \\
    FPF      & Fundamental Problem Formulation \\
\end{tabular}

\input{intro}
\input{graph_syntax}
\input{mcg_fpg}
\input{example_problem}

\section{Applications}

\section{Conclusions}

\bibliography{library}
\end{document}