### 2013-05-08 problem_formulation.tex

%\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
\documentclass[onecolumn]{svjour3} % insert '[draft]' option to show overfull boxes

 \title{A Graph Theory Based Approach to MDAO Problem Formulation}

\author{
  David J. Pate, %
  Justin Gray, %
  Brian J. German 
 }

 \institute { Justin Gray 
	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
		Brookpark Rd Clevland OH 44135 %
	\and David J. Pate 
	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
	\and Brian German 
	\at Assistant Professor, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.}

\journalname{Structrual and Multidisciplinary Optimization}

%\usepackage{setspace}
%\doublespace

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{appendix}
\usepackage[section]{placeins}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{booktabs}

\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}

\bibliographystyle{aiaa}

\usepackage{lettrine}
\usepackage{verbatim}

\newcommand{\txt}{\textrm}

%\usepackage{hyperref} %allows for the creation of actual text links
\begin{document}

\maketitle

\begin{abstract}
The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.  This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional design disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.  In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorically challenging.   The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.  In this paper, we propose an approach for addressing this problem based on the formalism of graph theory.  The approach begins by constructing the maximal connectivity graph (MCG) describing all possible interconnections between a set of analysis tools. Graph operations are then conducted to reduce the MCG to a fundamental problem graph (FPG) that describes the connectivity of analysis tools needed to solve a specified system-level design problem. The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.  The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.
\end{abstract}

\section*{Nomenclature}

\begin{tabular}{l l} 
    AAO      & All-At- \\
    MDAO     & Multidisciplinary Design Analysis and Optimization \\
    FPF      & Fundamental Problem Formulation \\
\end{tabular}

\input{intro}
\input{graph_syntax}
\input{mcg_fpg}
\input{example_problem}
\input{conclusions}

\bibliography{library}
\end{document}