### 2013-08-07 example_problem.tex

\section{Example Problem}
	\label{s:example problem}
	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. 
The example task is to create an FPG for the conceptual sizing of a single-aisle subsonic transport aircraft using a set of analysis tools with objectives including performance and gross weight for a required mission. 
	The full set of analysis codes available for use is provided in Table \ref{t:analysis codes}.
	\begin{table}[htbp]
	  \centering
	  \caption{Analysis tools for subsonic transport sizing}
		\begin{tabular}{cc}
		\toprule
		analysis tool & description \\
		\midrule
		VSP   & parametric vehicle geometry \\
		PDCYL & wing and fuselage weight estimation \\
		NPSS  & engine sizing and performance \\
		VORLAX & vortex lattice method  aerodynamics \\
		PMARC & low-order panel method  aerodynamics \\
		WATE  & engine weight estimation \\
		FLOPSa & mission performance \\
		  & engine sizing and weight estimation \\
		FLOPSb & mission performance only \\
		\bottomrule
		\end{tabular}%
	  \label{t:analysis codes}%
	\end{table}%
	Each analysis tool contributes individual disciplinary analysis capability, but the outputs of the tools are not mutually exclusive. 
	For example, VORLAX and PMARC are both aerodynamics codes that predict induced drag but with different levels of fidelity. 
	The Flight Optimization System (FLOPS) is included twice to represent two different configurations corresponding to different uses. 
	FLOPSa denotes FLOPS implemented to execute both mission performance and engine analysis, while FLOPSb indicates  FLOPS implemented to execute only mission performance analysis. 
	This representation is useful for ``supercodes'' that are capable of being implemented in different ways and with different sets of inputs and outputs.
The enumeration of these analysis tools and objectives concludes step (A) in Sec.~\ref{ss:process}.

Step (B) is the production of the MCG.
	In this example, the MCG is formed by using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis tool. 
	Some variables, such as geometry and performance, represent groups of variables that are passed as arrays or other data structures due to their similarity. 
	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
	\begin{table*}[htb!]
	  \centering
	  \caption{Analysis code input and output description}
		\begin{tabular}{ccccccccc}
		\toprule
		variable & \multicolumn{8}{c}{analysis code} \\
		\midrule
			  & VSP   & PDCYL & NPSS  & VORLAX & PMARC & WATE  & FLOPSa & FLOPSb \\
		geometry & in    & in    &       & in    & in    & in    & in    & in \\
		number of engines &       &       & in    &       &       &       & in    & in \\
		mission &       &       &       &       &       &       & in    & in \\
		fuselage weight &       & out   &       &       &       &       & in    & in \\
		wing weight &       & out   &       &       &       &       & in    & in \\
		engine weight &       &       &       &       &       & out   & out   & in \\
		wetted area & out   &       &       &       &       &       & in    & in \\
		inviscid drag &       &       &       & out   & out   &       &       & in \\
		drag  &       &       & in    &       &       &       & out   & out \\
		engine performance &       &       & out   &       &       & in    & out   & in \\
		performance &       &       &       &       &       &       & out   & out \\
		total weight &       & in    &       &       &       &       & out   & out \\
		\bottomrule
		\end{tabular}
	  \label{t:ins and outs}
	\end{table*}

	In this example, the MCG $M$ is formed in four steps:
	\begin{enumerate}
	\item An analysis block is created for each analysis code using the information in each column of Table \ref{t:ins and outs}. 
	Each analysis block is formed by first creating variable nodes for each input and adding directed edges into a single function node. 
A function edge is created from this function node into a second function node, which is then directed into variable nodes corresponding to each output. 
	A sample analysis block is shown for analysis code FLOPSa in Fig.~\ref{f:FLOPSb analysis block}.
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=3.25in]{images/FLOPSa_analysis_block}
	  \end{center}
	  \caption{Sample analysis block for analysis code FLOPSa using Table \ref{t:ins and outs}}
	\label{f:FLOPSb analysis block}
	\end{figure}

	\item Expression blocks are created to represent the objectives for performance and total weight.

	\item A variable node is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).

	\item Connection edges are created that connect each variable node to every other variable node representing variables with the same name. 
	The direction is determined by whether the variable node has an edge directed into or out of a function node, i.e. whether it is a local input or a local output.
	\end{enumerate}
	The resulting MCG is shown in Fig.~\ref{f:MCG holes}. 
	Cycles are shown as dashed lines. FPGs formed from this MCG may or may not retain these cycles,  depending on user choices.
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=6in]{images/MCG_edit_holes}
	  \end{center}
	  \caption{Maximal connectivity graph for the subsonic transport example problem}
	\label{f:MCG holes}
	\end{figure*}

To implement step (C) in Sec.~\ref{ss:process}, the indegree of every variable node is set to unity to reflect a single fidelity analysis for which the design variables have not yet been selected.

Step (D) now procededs by executing the FPG algorithm. 
An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb. 
These holes are propagated upstream by the FPG algorithm and create holes in the objectives, thereby preventing a valid FPG.

Step (D)(a) begins to resolve this conflict by creating an input for the number of engines. 
In this example, the number of engines was not originally included as an input to demonstrate how the process can reveal the need to select a variable as an input.
Step (D)(b) then reduces the lower and upper indegree limits for the variable node representing the mission definition to zero, implying that these nodes must be specified as design variables (see Table \ref{t:variable node classification}). 
The new MCG is shown in Fig.~\ref{f:MCG}.
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=6in]{images/MCG_edit}
	  \end{center}
	  \caption{The updated MCG with a new input.}
	\label{f:MCG}
	\end{figure*}

	Although an FPG may now be obtained, this process has not provided a method to decide which edges to retain when resolving a collision, and these decisions will likely determine which analysis tools are included in the FPG. These decisions are left to the discretion of the designer in the context of the specific implementation. 
	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analysis. 
	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 
There are two cycles which arise from each local output from PDCYL being directed into FLOPSa and then back into PDCYL.

	Similar alternatives would be to minimize the number of analysis blocks involved in cycles, counting multiplicity, or to minimize the length of the longest cycle (called the \emph{circumference} of the graph). It is beyond the scope of this work to explore efficient methods for finding the FPGs meeting these criteria. However, as discussed by Gabow \cite{Gabow1985}, graph algorithms typically scale well with the size of the graph. This suggests that the computation cost of finding an FPG should be small even for large problems.
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=4.5in]{images/FPG_edit_fewest_cycles}
	  \end{center}
	  \caption{FPG with the fewest number of cycles.}
	\label{f:FPG fewest cycles}
	\end{figure*}

	Alternatively, it may be desirable to resolve collisions by preferentially choosing analysis tools with favorable properties, such as higher accuracy or minimal run time, whenever possible.
	One method to implement an FPG in this paradigm is to use a ranking system. Each analysis block is assigned a value by the designer describing how desireable it is for this block to be present in the FPG. The connection edges directed out of each analysis block are then assigned a weight equal to the value assigned to the analysis block. Finally, each collision is resolved by selecting the edges with the highest weights.
For the current example problem, consider the rankings for each analysis code given in Table \ref{t:rankings}.
	\begin{table}[htbp]
	  \centering
	  \caption{Example ranking of importance}
		\begin{tabular}{cc}
		\toprule
		analysis code & rank \\
		\midrule
		VSP   & 5 \\
		PDCYL & 5 \\
		NPSS  & 4 \\
		PMARC & 4 \\
		FLOPSb & 4 \\
		VORLAX & 3 \\
		WATE  & 3 \\
		FLOPSa & 2 \\
		\bottomrule
		\end{tabular}%
	  \label{t:rankings}%
	\end{table}%
	The resulting FPG has four cycles and is shown in Fig.~\ref{f:FPG highest rank}. 
	\begin{figure*}[htb!]
	  \begin{center}
		\includegraphics[width=6in]{images/FPG_edit_ranking}
	  \end{center}
	  \caption{FPG obtained using a ranking system.}
	\label{f:FPG highest rank}
	\end{figure*}

	Finally, consider a case in which 
the induced drag input into FLOPSb is a multi-fidelity input, meaning that 
 multiple analysis codes calculate the same variable. 
	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repreating the (automated) process. 
	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.